### 2013-03-27 doc/final_report/implementation.tex
Finished paper preprocessing section in testing and started writing stuff for web backend

@@

\subsection{Pre-processing with SAPIENTA}

@@

SAPIENTA for Python also relies upon a package called CRFSuite which implements
Conditional Random Fields, a method for segmenting and labelling sequence
data\cite{CRFsuite}. This library did not compile properly on the test
environment and its creator had to be contacted via a mailing list (See
Appendix \ref{sec:crfemail}. After a few days, the owner responded and the
library was compiled successfully.

@@

\subsubsection{ Feature Selection and Clustering}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=0.6\textwidth]{images/implementation/review_corescs.png}
\caption{Contrasting CoreSC content of Review and Research Papers}
\label{fig:coresc_pies}
\end{figure}

Figure \ref{fig:coresc_pies} shows the CoreSC content of a review aper and a
research paper randomly selected from the corpus. The review papers tend to be
made up almost entirely from Background CoreSC sentences. However, research
papers are much more evenly spread, made up of several different types of
CoreSC. This investigation suggested that there is almost certainly a discriminative
relationship between CoreSC categories and a paper's type.

Although all of the CoreSC types could be influential in determining a paper's
type, it was likely that some were more important than others. A further
investigation was carried out to discover the most effective features for
categorising the data. Since no classifier model had been trained at this
point, the data was clustered into three unlabelled groups using the K-means
clustering algorithm. This algorithm was run with three CoreSC

@@

@@

@@

@@

During the development of Partridge, the structure of the SQL database used to
store the papers changed a lot due to modifications to the ORM modules as
discussed above. These structural changes meant that the production database at
\url{http://farnsworth.papro.org.uk/} and the development database were not
compatible and that updates to the code on the development server would break
the application on the prodution server unless the database structure was also
updated.

@@

>>>>

Partridge's Paper Preprocessor module was the most difficult part of the
project to implement. From acquiring papers for use in the project to teaching
a machine learning system to recognise different types of scientific
publication, a number of technical challenges were encountered.

@@

\subsection{Annotating with SAPIENTA}

@@

SAPIENTA for Python also relies upon a package called CRFSuite which implements
Conditional Random Fields, a method for segmenting and labelling sequence
data\cite{CRFsuite}. This library did not compile properly on the test
environment and its creator had to be contacted via a mailing list (See
Appendix \ref{sec:crfemail}). After a few days, the owner responded and the
library was compiled successfully.

@@

\subsubsection{Feature Selection and CoreSCs}

@@

\begin{figure}[!h]
\centering
\begin{subfigure}[b]{0.6\textwidth}
\includegraphics[width=\textwidth]{images/implementation/review_corescs.png}
\caption{CoreSC content for a sample review paper}
\end{subfigure}
\begin{subfigure}[b]{0.6\textwidth}
\includegraphics[width=\textwidth]{images/implementation/report_corescs.png}
\caption{CoreSC content for a sample research paper}
\end{subfigure}

\caption{Contrasting CoreSC content of Review and Research Papers}
\label{fig:coresc_pies}
\end{figure}

Figure \ref{fig:coresc_pies} shows the CoreSC content of a review paper and a
research paper randomly selected from the corpus. The review papers tend to be
made up almost entirely from Background CoreSC sentences. However, research
papers are much more evenly spread, made up of several different types of
CoreSC. This investigation suggested that there is almost certainly a discriminative
relationship between CoreSC categories and a paper's type.

\subsubsection{ K-means clustering}

Before working with supervised learning systems, a preliminary experiment was
carried out to see whether clustering papers using the K-means algorithm could
successfully identify the paper types, thus removing the need to train and
maintain a supervised model. This was also an excellent opportunity to explore
the data and try and detect any inherent patterns within the papers.

K-means clustering uses the Euclidean distance between each data sample to
decide which cluster it is closest to and should belong to. K is the number of
clusters that are used and is decided ahead of execution. K-means clustering is
often run multiple times using a range of K values when the number of clusters
in the data is unknown. The performance of the clustering algorithm is often
evaluated using the K-means silhouette which is a measurement of how tightly
the data is clustered. A low silhouette value ($K \approx 0$) suggests loose
clustering and therefore little correlation, a high silhouette value ($K
\approx 1$) corresponds with very tight clustering and good correlation.

All of the papers were manually identified as either a ``Case Study", a
``Review Paper" or a ``Research Paper." They were then loaded into a script and
clustered using their CoreSC respective compositions. K-Means silhouettes were
used as a measure of how effective the clustering process was. This was
repeated for $ K \in [2..8] $ and the best silhouette measurement and value for
K retained. The algorithm consistently showed that K=3 was the best K value.
However, the silhouette value was consistently low. It was suggested that this
may have been because all CoreSC types were considered equally; the less
discriminative CoreSCs may have been offsetting the effect of the highly
discriminative ones.

To remedy this, all papers were again clustered, but using only three CoreSC
labels at a time: reducing the noise from the less important types. The
algorithm was run for all combinations of 3 CoreSC labels, again retaining only
the most successful combination of features and K value based upon the K
silhouette value. This returned a set of much stronger K silhouette values in
general. The algorithm also determined that the combination of ``Background",
``Motivation" and ``Method" CoreSC labels seemed to be the most
discriminative, providing a silhouette value of 0.63.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{implementation/cluster.png}
\caption{Paper types rendered using Bac, Met and Mot CoreSCs}
\label{fig:clusters}
\end{figure}

Figure \ref{fig:clusters} shows the papers from the corpus rendered in 3
dimensions using their Background, Method and Motivation CoreSC makeup.
Although the data does not cluster perfectly by class, these three categories
do provide fairly clear discrimination between the three types of paper.

Clustering the papers did not seem to provide a reliable way to classify them
by type and it was decided that better results could be obtained through the
use of a supervised learning classifier instead. However, the three most
discriminative CoreSC types were chosen as features for the learner.

\subsubsection{Random Forest Learner}

To classify the papers, a Random Forest Learner was trained using a subset of
the Partridge corpus using Background, Method and Motivation CoreSC information
from the papers. The classifier was found to be highly accurate, correctly
predicting a paper's type in 87.9\% of cases. The classifier's accuracy and
evaluation are further discussed in Section \ref{sec:evaluation_learners}
below.

The new classifier was incorporated into the paper preprocessor and triggered
after SAPIENTA annotation for each new paper. It was also called
retrospectively for papers in the Farnsworth Partridge instance that were not yet
classified.

@@

Providing notifications to users when a paper has been processed was not part
of the original design for Partridge. The feature was added to the system after
being raised as a potential improvement by one of the test users. Previously,
users to added papers to the Partridge corpus did not know whether the
operation was a success unless they ran a search and the paper appeared in
their results. Adding email notifications to Partridge significantly improved
the user experience for paper uploaders. The uploader is given the option to
provide their email address when they send the paper to the Partridge server.
Once the preprocessor has converted, annotated and classified a paper, an email
is sent to the uploader informing them that the paper has been added to the
database and containing a direct link to the paper's profile page.

The system also emails users upon failure, allowing them to report the error to
the GitHub page if required and also providing closure for users who are unsure
why their paper isn't appearing in Partridge queries.

@@

The Partridge web backend was much easier to implement than the preprocessor.
The web framework Flask was used to do a lot of the hard work such as routing
requests and encoding data into browser-compatible formats.

@@

Flask provides a model-view-controller style framework in which specific web
URLS are represented by `view' functions.

@@

During the development of Partridge, the structure of the SQL database used to
store the papers changed a lot due to modifications to the ORM modules as
discussed above. These structural changes meant that the production database at
\url{http://farnsworth.papro.org.uk/} and the development database were not
compatible and that updates to the code on the development server would break
the application on the production server unless the database structure was also
updated.

To prevent this from happening, a Python utility called
Alembic\cite{alembic2013} was used for keeping the Database structure under
Version Control. Alembic is executed when Partridge's ORM models are modified
and can detect differences in database structure between the SQLAlchemy model
in Partridge's source code and the live database. It generates sets of SQL
commands that are used to synchronise the database structure with the most
recent program structure. This allowed the Farnsworth database to be updated
automatically when a new change to Partridge's data structures was made rather
than having to be purge and repopulate it.

@@

Flask provides a test server that can be used for developing your web
application. However, it is written in python and is not designed to be as
stable or scalable as a full web server application such as Apache. Once
Partridge's basic web interface was complete and the application ready to be
deployed, it had to be integrated into Apache to provide stable web server
support.

By default, Apache only serves static HTML content and image/data files. It makes
use of plugins to support dynamic web pages, in languages such as PHP and
Python. A plugin called WSGI is used to allow Apache to serve Python
applications. WSGI requires that the Python program exposes an
\emph{application} object which implements a common interface in order to
communicate with Apache.

After thorough investigation, it was discovered that Flask provides a WSGI
compatible \emph{application} object. A small wrapper was written that exposes
the Flask app to WSGI and the server began to work as expected. Unfortunately,
the WSGI interface didn't allow the preprocessor daemon to be run in the
background. Therefore, a second wrapper was created that allowed the execution
of the Paper daemon as a standalone process.
